---
title: "R and RMarkdown Intro"
author: "Amanda B. Young"
date: "10/8/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Header 1
**bold text**
*italics*

You can just type free text
this will be on the same line

This is a new paragraph

1. apple
1. bannana
1. orange


1. apple
1. bannana
1. orange

```{r, first_chunk}
# comment
x<-12.12345
```
Tthe result is `r round(x,2)`

## Random
### This code sets up a random distribution of 100

$$\sum{x^2}$$

```{r second chunk}
x<-rnorm(100)
hist(x)
```

THis is the mean of the random distribution `r mean(x)`

```{r}
x<-c(1,2,3)
mean_x<-mean(x)
```

# Read in a file
Read in a file downloaded from the Arctic Data Center. This is from [Tweedie 2009](https://doi.org/10.18739/A25T3FZ8X)

```{r}
bg_chem<-read.csv(file = "BGchem2008data.csv" , stringsAsFactors = FALSE)
head(bg_chem)
```

`$` list selector operator to select a particular column

Short cut to run a chunk is Ctrl+Alt+C

```{r}
mean_temp <- mean(bg_chem$CTD_Temperature)
mean_temp
```

```{r}
plot(x = bg_chem$CTD_Depth, y=bg_chem$CTD_Temperature)
```

