---
title: "Data-tidying-lesson"
author: "Amanda B. Young"
date: "10/9/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, warning = FALSE, message= FALSE}
library(tidyr)
library(ggplot2)
library(dplyr)
```

To access `filter` from the `stats` package: `stats::filter()`

#Read in and clean up data

```{r open file}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)

```

`stringsAsFactors` reads characters as factors rather than characters, keeps the future issues down
Try to allways set `stingsAsFactors` to `FALSE`


Select the columns we want using the `select()`

```{r select-columns}
catch_data<-catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) 
  select(-All, -notesRegCode)
head(catch_data)

```

```{r summary-data1}
summary(catch_data)
```

Change the values in the CHinook column to numeric using `mutate()`
```{r clean-data1}
catch_clean<- catch_data %>% 
  mutate(Chinook=as.numeric(Chinook))
head(catch_clean)
```

Investigate using `which()` and `is.na()`

```{r NA-check}
i<- which(is.na(catch_clean$Chinook))
i
```
```{r}
catch_original[i,]
```

* use `mutate` to change the I to 1
* use `mutate` to coerce Chinook column to numeric
` %in%` can be used instead of `|` as an or when using ifelse
```{r keep-cleaning}
catch_clean<-catch_data %>% 
  #mutate(Chinook = ifelse(Chinook == "I"| Chinook =="l")) %>% 
  mutate(Chinook = ifelse(Chinook  %in%  c("I", "l"), 1, Chinook)) %>% 
  mutate(Chinook=as.numeric(Chinook))
head(catch_clean)
```

# Tidy Data
Move from wide formate to tall format using `pivot_longer`
```{r long}
catch_long<-catch_clean %>% 
  pivot_longer(cols=-c(Region, Year), 
               names_to = "Species",
               values_to = "Catch")
head(catch_long)
```

```{r wide}
catch_wide<-catch_long %>% 
  pivot_wider(names_from = Species,
              values_from = Catch)
head(catch_wide)
```

`rename()` catch to catch_thousands
`mutate()` to 
```{r thousand}
catch_long<-catch_long %>% 
  mutate(Catch=Catch*1000) 
```

# Summarize Data
mean catch by region of Sockeye

```{r fiter-group-Sockeye}
mean_region<-catch_long %>% 
  group_by(Region) %>% 
    filter(Species == "Sockeye") %>% 
summarise(mean_catch=mean(Catch),
            n_obs=n())
head(mean_region)
```

total annual catch
```{r}
annual_catch<-catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(Catch)) %>% 
  arrange(desc(total_catch))
annual_catch

```

```{r}
ggplot(annual_catch, aes(x=Year, y=total_catch))+
  geom_point()
```

```{r}

catch_long %>% 
  group_by(Year,Species) %>% 
  summarise(total_catch = sum(Catch)) %>% 
  arrange(desc(total_catch)) %>% 
  ggplot(aes(x=Year, y=total_catch, colour=Species))+
  geom_point()

```
)

```{r regions}
region_defs<- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"),
                    stringsAsFactors = FALSE) %>% 
  select(code, mgmtArea) %>% 
  rename(Region = code)
head(region_defs)


```

Create a join between catch and region files
```{r join}
#catch_joined<-left_join(catch_long, region_defs, by=c("Region"= "code"))
catch_joined<-left_join(catch_long, region_defs, by="Region")
catch_joined<- left_join(catch_long, region_defs)
head(catch_joined)
```

```{r seperate}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE) %>% 
  unite(date_2, year, month, day, sep="-")
```

can use substr to seperate a variable by the number of integers, also can use package `stringr::str_sub` to get the last few integers
`string_pad` can say I want two digits but if I only have a 5 then it would become 05

```{r}

```

